---
title: "MA 415/ MA615 
        HW 3 Section 1 project"
output: ioslides_presentation
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Team members|Group 6

- Yue Wu 
- YinChia Huang 
- Francesco Ignazio Re

## Outline

- Introduction
- Data set Visualization & Analysis
- Statistic Analysis
- Time Series & Analysis
- Conclusion

## Introduction

Apprehensions at the US-Mexico border have declined to near-historic lows over the last few years.
	The objective of this report is to give a deeper insight on this change that has been occuring.
	Through the analysis of the data collected by the U.S. Customs and Border Protection through the years,
	we intend to shed light on the general trend of this phenomenon, focusing on how factors such as time and place have influenced the given outcome.

```{r include=FALSE}
# Raw Data
PB2010 <- read.csv("PB2010.csv")
PB2017 <- read.csv("PB2017.csv")
```

## Overview of 2010 Apprehensions 
The gray dots are Sector "Tucson", which had the most unstable number of apprehensions and was the sector who contributed the most in the apprehensions of 2010. 
```{r}
plot(PB2010$Month, PB2010$Apprehensions, type = "p",col = PB2010$Sector, main = "Apprehensions by month in 2010",ylab= "Apprehensions", xlab= "Month" )
legend("topright",inset=c(0.22,0),
       c("Big Bend","Del Rio", "El Centro","El Paso","Laredo","Rio Grande Valley", "San Diego", "Tucson","Yuma"),
       cex=0.8,
       bty="n",
       fill=c("gray0","red","green","blue","cyan", "magenta","yellow","gray", "black"))


#  install.packages("tidyverse")
#  library(tidyverse)
# #
#  ggplot(data = PB2010) + 
#    geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))

```

## Overview of 2017 Apprehensions
The pink dots are Sector "Rio Grande Valley" which contributed the most in the apprehensions of 2017. Notably,apprehensions dramatically increased from Sep to Oct.

```{r}
plot(PB2017$Month, PB2017$Apprehensions, col = PB2017$Sector, main = "Apprehensions by month in 2017",ylab= "Apprehensions", xlab= "Month")

legend("topleft",inset=c(0.05,0),
       c("Big Bend","Del Rio", "El Centro","El Paso","Laredo","Rio Grande Valley", "San Diego", "Tucson","Yuma"),
       cex=0.8,
       bty="n",
       fill=c("gray0","red","green","blue","cyan", "magenta","yellow","gray", "black"))

# install.packages("tidyverse")
# library(tidyverse)
# 
# ggplot(data = PB2017) + 
#   geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))

```


```{r include=FALSE}
#reading and setting up the 2010 apprehension data set
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-1]
rownames(ap10) <- cities
ap10 <- cbind(ap10[4:12],ap10[1:3])
SumCol10 <- colSums(Filter(is.numeric, ap10))

#reading and setting up the 2017 apprehension data set
ap17 <- read.csv("BP Apprehensions 2017_Windows.csv")
ap17 <- ap17[-c(1,2,15)]
ap17 <- cbind(ap17[4:12],ap17[1:3])
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))

#Creating matrixes needed for the bar plot graphs:

#1) Creating a matrix containing the Sum of the Columns of each data set
CmbColS <- rbind((SumCol10),(SumCol17))
#2) Creating a matrix containing the Sum of the Rows of each data set 
CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))

#reading and setting up the 17-years apprehenesions data set
msum  <- read.csv("BP monthly summaries.csv", stringsAsFactors = FALSE)
years <- msum[c(1)]
msum  <- msum[-c(1)]
rownames(msum) <- as.vector(t(years))
msum
msum <- cbind(msum[4:12], msum[1:3])
msum<-msum[18:1,]
ts3 <- ts(as.vector(unlist(t(msum))),frequency = 12, start=c(2000,1))


# 3 MONTHS PEAK 2010
#Testing the hyphotesis for which the three most trafficated months'observations in 2010 are equal in mean to the 
#same months in 2017
x10 <- msum[11,]
x17 <- msum[18,]
xa10 <- x10[10:12]
xa17 <- x17[10:12]


# 3 MONTHS PEAK 2017
#Testing the hyphotesis for which the three most trafficated months'observations in 2017 are equal in mean to the 
#same months in 2010

xb17 <- x17[3:5]
xb10 <- x10[3:5]



```


## 2010 v.s 2017 Apprehensions by month  
The most trafficated months in 2010, such as March, April and May are also the ones with the biggest decline in 2017. 
```{r echo=FALSE}
barplot(CmbColS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by month",
        cex.lab =1.5,
        col = c("orange","red") )
legend("topright",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=c("orange","red") )

```

## 2010 v.s 2017 Apprehensions by sector 
The greatest change has occured in Tuscon, the area with the highest number of apprehensions in the 2010, that observed a drop of over the 80% according to the data collected in 2017.
```{r echo=FALSE}
barplot(CmbRowS,
        beside=TRUE, 
        las=2, main = "Apprehensions by sector", 
        cex.lab =1.5, 
        col = c("orange","red") )
legend("topleft",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=c("orange","red") )


```

## Descriptive Data Analysis 
From 2010 to 2017, the U.S. Customs and Border Protection saw an overall 36 percent decrease in apprehensions for illegal entry to the country. The plots show a significant different trend of monthly apprehension changes between the two years. Sectors also exhibit different patterns of changes from each other.

## Statistic Analysis(1)
###Let's use simple statistical tests to compare the change witnessed in the the sector with the most apprehensions in 2010 and 2017.

--- 

#### *2010 max-sum sector: No.8 Tuscon*
```{r, echo=FALSE}

#function to get t-tests for by-sector comparison
by_sector.t <- function(y){
  
  t.test(as.numeric(ap10[y,]), as.numeric(ap17[y,]), paired = TRUE)
}

# T-test on the sector 8 Tucson (The most trafficated sector in 2010)
by_sector.t(8)

```
######**t test:** p= 6.324e-05 < 0.05  
######**Conclusion:** We can reject the null hypothesis that Tuscon has equal means in 2017 from 2010.  
######**Interpretation:** Tuscon does  play a critical role in the fluctuations of apprehensions between 2010 and in 2017.  

---

#### *2017 max-sum sector: No.6 Rio Grande Valley*
```{r, echo=FALSE}
# sector with max yearly sum in 2017 --> test on max-apprehension sector changes
by_sector.t(6)
```
######**t test:** p= 0.03167 < 0.05  
######**Conclusion:** We can reject the null hypothesis. RGV has different means in 2010 from 2017, with 95% confidence.  
######**Interpretation:** RGV plays a critical role in the fluctuations of apprehensions between in 2010 and in 2017, however not as much as the the sector Tuscon.

## Statistic Analysis (2) 
###Testing the hyphotesis for which the three months with apprehension peak in one year are equal in mean of the same months in the other. 

---

#### *Peak in 2010: March, April, May*
```{r, echo=FALSE}


t.test(as.numeric(xb10), as.numeric(xb17), paired = TRUE)
```
######**t test:** p= 0.014 < 0.05  
######**Conclusion:** We can reject the null hypothesis. March - May 2010 means are significantly different from those in 2017, with 95% confidence.  
######**Interpretation:** The peak in 2010 is significantly different in value from the same months in 2017.     

---

#### *Peak in 2017: October, November, December*
```{r, echo=FALSE}

t.test(as.numeric(xa10), as.numeric(xa17), paired = TRUE)

```
######**t test:** p= 0.081 < 0.05  
######**Conclusion:** We cannot reject the null hypothesis. Oct - Dec 2017 means are not significantly different from those in 2010.  
######**Interpretation:** The peak in 2017 is not significantly different in value from the same months in 2010. The test result correspond with the narrow boxplot trend during those months. However, the small degrees of freedom makes the test less reliable. 

## Time series graph{.smaller} 

The time series plot shows that from 2000, there has been continuous decline, from a high of over 1.6 million in 2000 to around 300,000 in 2017. Over these years, US policies on immigration control have been rapidly developed, leading to potential correlation with the change in apprehensions.


```{r echo=FALSE}
#Plot the time Series Graph
ts.plot(ts3,main="Time Series from 2010 to 2017", gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)))

#create data for calculating the annual means and draw the means line
ts3.1<-as.vector(ts3)


for (i in 1:18 ){
  if (i == 1){
    from <- i
    to <- 12
    annualMeans <- c(mean(ts3.1[from :to]))
  }  
  else{
    from <- 1 + to
    to <- i*12
  }  
  annualMean<- mean(ts3.1[from: to])
  #print(annualMean)
  annualMeans <- append(annualMeans, annualMean)
  segments(1999+i,annualMean,2000+i,annualMean, col = "blue",lwd=2)
}
annualMeans <- annualMeans[-1]  

## add text labels (year)
text(0.5+seq(from = 2000, to = 2017, by = 1),annualMeans[1:18]+6000,labels=2000:2017, cex=0.7, col = rgb(.9,  0, .7, .5) , font=2)
## add legend
legend('topright', col=c(rgb(0,0,0),"blue"), lty=1, lwd=2, 
       legend=c("Monthly apprehensions", "Annual averages"), bg='white')


```

## Time series analysis (1)
Box plot across months will give us a sense on seasonal effect. We can see that the number of apprehension in March has the biggest range and highest median.

```{r echo=FALSE}
# Box plot across months will give us a sense on seasonal effect 
boxplot(ts3~cycle(ts3),ylab= "Apprehensions", xlab= "Month", main="Monthly Box plot " )
```


##Time series analysis (2){.smaller}
In addition to the overall decline, seasonal fluctuation is another noticeable trend in the plots. A reasonable explanation is that harsh weather decreases attempts of illegal entry while keeping other factors controlled. The seasonal effects cause the changes among months, but plays little role in the rapid changes among years.

```{r echo=FALSE}
# Decompostion of multiplicative time series
timeseriescomponents <- decompose(ts3, type="multiplicative")

#timeseriescomponents
plot(timeseriescomponents)

```


## Time series analysis (3)
Month plot graph displays the time series plot for each month from 2000 to 2017. The shape of each month is very similar with different magnitudes.  
```{r}
monthplot(ts3, ylab= "Apprehensions", xlab= "Month", main="Monthly plot with monthly mean")
```

## Time series prediction
Using ARIMA Model to predict future apprehensions
```{r}
fit <- arima(ts3, order=c(1,0,0), list(order=c(2,1,0), period=12))
fore <- predict(fit, n.ahead=24)

#  error bounds at 95% confidence level
U <- fore$pred + 2*fore$se
L <- fore$pred - 2*fore$se
ts.plot(ts3, fore$pred, U, L, col=c(1,2,4,4), lty = c(1,1,2,2))
legend("topleft", c("Actual", "Forecast", "Error Bounds (95% Confidence)"), col=c(1,2,4), lty=c(1,1,2))

```

## Conclusions

The significant decline of border apprehensions since 2000 mainly come from changes in specific sectors, such as Rio Grande Valley, and time of a year, such as March to May. More research on political and economical factors can be done to further explain the causes.




## Reference
[Using R for Time Series Analysis](https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html#time-series-analysis)
<br> [Time series decomposition](http://www2.hawaii.edu/~fuleky/econ427/6_Time_series_decomposition.html#1) 
<br> [Time Series Analysis in R](https://rstudio-pubs-static.s3.amazonaws.com/254962_6d87741d9ad64a32b6e47d684875dde9.html) 
<br> [Introduction to Forecasting with ARIMA in R](https://www.datascience.com/blog/introduction-to-forecasting-with-arima-in-r-learn-data-science-tutorials) 
<br> [Border apprehensions hit 17-year low in March](http://www.washingtonexaminer.com/border-apprehensions-hit-17-year-low-in-march/article/2619338)


